\section{Introduction} \label{sec:introduction}

Software being a ubiquitous technology, it is nowadays being used in a wide spectrum of domains: music, teaching, arts, engineering.
The proliferation of visual \acfp{ide} has helped better assist these domain users who are unfamiliar with programming.
Common requirements to \acp{ide} in general often include~\cite{habermann1986}: uniformity and consistency across the different tools it provides, an interactive user interface for performing the different tasks, ability to inspect a well-defined state of the system being developed (\eg debugging), integrate control of versions of the system in an individual or collaborative environment, and ensure a manageable development process.

Previous studies have surveyed \acp{ide}, but they were mostly focused on the
programming domain. \cite{hils1992, fischer1994, habermann1986}
\es{cite}\jmr{addressed}
In this study, we focus on \acp{ide} used in non-programming domains where the underlying language is visual (\eg diagrams), as opposed to textual (\eg source code).
Often these \acp{ide} provide the user with information about the syntax of the supported language(s) or otherwise provide features that integrate with any constraints of the visual language.
Components and behaviors of the \ac{ide} interface can greatly affect the overall usability of an \ac{ide} as well as its suitability to its supported language.
Consequently, it is important to utilize as many techniques as possible to ensure both overall interface usability and suitability to the target language.
Several works have looked at metrics for measuring the usability of
software. \cite{constantine1996}
\es{cite}\jmr{addressed}
We have adapted these metrics to specifically look at visual \acp{ide} in different domains.

\Sect{sec:methods} describes the methodology and features covered by our survey.
\Sect{sec:results} summarizes the results of the study for each \acp{ide} surveyed.
In \Sect{sec:discussion}, we discuss interpretation and limitations of the survey and we conclude in \Sect{sec:conclusion}.

%A good starting point in the analysis of visual language \ac{ide}
%suitability is an understanding of the visual languages themselves. The
%interface supporting a language must ensure that the design choices made
%within the language are mirrored and supported within the \ac{ide}.
%\cite{hils1992} For example, languages with a high level of liveness
%\cite{hils1992} would not be well suited to \acp{ide} developed with an
%indirect trigger output generation style as any benefits from liveness are
%lost. Additionally, domain-specific features of domain-customized languages
%may introduce \ac{ide} requirements to maintain full support.
