\section{Methods} \label{sec:methods}

The work began with the selection of a variety of visual language \acp{ide} from
differing domains of use. This was followed by the creation and definition
of a set of novel visual \ac{ide} interface features. Finally, each of the
selected \acp{ide} were evaluated based on the set of interface features.
\Sect{subsec:ideselection} describes the process and criteria for \ac{ide}
selection as well as the specific target \acp{ide} selected for study.
\Sect{subsec:featuredefinition} details the set of novel interface features
devised to study the selected visual \acp{ide}, and \Sect{subsec:ideevaluation}
describes the process of evaluation taken for each \ac{ide}.


\subsection{\acs{ide} Selection} \label{subsec:ideselection}

\ac{ide} selection was a non-scientific process with three primary criteria
driving selection. First, \acp{ide} were selected on the basis of their support
for a visual language as the primary language of development. All \acp{ide}
studied therefore are primarily visual language \acp{ide}. Subsidiary criteria
for selection included popularity of the \ac{ide} within its domain as well as
representation from a variety of development domains. One goal of this
study was the evaluation of visual \acp{ide} from a number of different domains,
so \acp{ide} from eight independent domains were selected. The specific domains
under consideration included 3D modeling, animation, modeling, music,
prototyping, simulation, software, and workflow.


\subsection{Feature Definition} \label{subsec:featuredefinition}

Through observation and use of the \acp{ide}, we began noting the differences
and similarities present that defined either interface purpose or
functionality. We determined whether these interface features could be
applied to all \acp{ide} and subsequently formalized the definitions of all
universally applicable features. Some variables were strongly influenced by
existing literature. All variables were categorized under five headings:
Audience, Chrome, Human Interface, Integration, and Language Syntax.
\Sect{subsubsec:audience} --- \Sect{subsubsec:languagesyntax} detail these
categories and constituent features.


\subsubsection{Audience Features} \label{subsubsec:audience}

The audience of an \ac{ide} refers generally to the target populations for which
the \ac{ide} is developed or intended. Features under the audience category
attempt to describe those meta-features of \acp{ide}.


\paragraph{Domain} The domain of an \ac{ide} refers to the domain or field of
knowledge under which the interface falls. That is, the field or fields for
which the interface is primarily used, if any. This is a nominal variable.
Example values might include \texttt{General}, \texttt{Modeling},
\texttt{Software}. A value of \texttt{General} indicates that the \ac{ide} is
general purpose and can be applied to or used within any number of
different fields to the same extent.


\paragraph{Skill Level} The skill level of an \ac{ide} describes what level of
skill in the \ac{ide}'s domain is expected of users. This is a nominal variable.
Possible values are \texttt{Novice}, \texttt{Intermediate},
\texttt{Expert}, and \texttt{General} with a value of \texttt{General}
indicating that the \ac{ide} offers a powerful set of advanced features while
maintaining components that emphasize accessibility and ease of use.


\subsubsection{Chrome Features} \label{subsubsec:chrome}

The chrome of an \ac{ide} is the total set of all user interface components
external to the workspace. This includes every tool, menu, button, or other
user interface component not contained within the workspace area.


\paragraph{Popular Features} Many \acp{ide} provide support for a specific
subset of common operations. \citeauthor{murphy2006} defines a list of the
top ten \ac{ide} features most frequently utilized by developers which is
reproduced here as a set of boolean sub-variables. Each sub-variable
indicates whether the \ac{ide} in question supports the use of that particular
features. This is a compound variable.

\subparagraph{Delete} Delete a syntactic element in the workspace.

\subparagraph{Save} Save and export a model to storage media.

\subparagraph{Paste} Duplicate an existing syntactic element in the
work{-}space from the paste buffer.

\subparagraph{Content Assist} Provide suggestions or completion for
elements.

\subparagraph{Copy} Place an existing syntactic element from the
work{-}space into paste buffer.

\subparagraph{Undo} Undo the user's most recent action.

\subparagraph{Cut} Remove a syntactic element from the workspace and place
it in the paste buffer.

\subparagraph{Refresh} Load contents of workspace and interface dashboard
elements from storage media and update display if necessary.

\subparagraph{Show View} Open and display a new tool in the interface.

\subparagraph{Next Word} Move active selection to the next element
according to some natural ordering.


\paragraph{Context Sensitive Tools} Any interface component which changes
visibly or is generated anew depending on the context of selected elements
within the workspace is context sensitive. This is boolean variable
indicates whether context sensitive tools are supported.


\paragraph{Degree of Interface Visual Richness} The degree of an \ac{ide}'s
interface visual richness describes the extent to which it utilizes visual
variables to increase the visual discriminability of available tools. This
is a compound variable composed of eight sub-variables. Each sub-variable
is a boolean measure determining whether or not the described visual
variable is utilized in the interface to distinguish available tools.

\subparagraph{Icons} Images contained in a border of a standard size and
shape representing distinct actions or tools.
\cite{costagliola2002,moody2009}

\subparagraph{Shape} Distinct shapes indicate different tools.
\cite{moody2009}

\subparagraph{Size} Different tool sizes indicate distinct tools.
\cite{moody2009}

\subparagraph{Color} Color is used to indicate distinct tools.
\cite{moody2009}

\subparagraph{Text} Text (or typographic variation) is used to identify or
distinguish tools.~\cite{moody2009}

\subparagraph{Organizational Coherence} Components with related purpose are
visually grouped in the interface.~\cite{constantine1996}

\subparagraph{Texture} Shading or shadows are used to modify tools or to
distinguish between distinct tools.~\cite{moody2009}

\subparagraph{Brightness} The brightness of a color (\ie its perceived
luminosity) is used to indicate a difference between tools.
\cite{moody2009}


\paragraph{Multiplicity of Perspectives} This ratio variable encapsulates
the number of available predefined interface perspectives available to the
user. A perspective is defined as a visual configuration of the available
tools in the interface chrome and elements in the workspace for the
purposes of accomplishing a distinct task by mean of a distinct process.
Values can range anywhere greater than zero.


\paragraph{Object Properties Window} The object properties window is any
interface component that displays the properties of an element in the
workspace. This component generally also allows modification of element
properties. This is a nominal variable. If no object properties window is
available, this variable takes the value \texttt{None}. Otherwise,
available values are \texttt{Omnipresent} and \texttt{Manual}. The former
refers to a window which is always present, allowing contents to update
contextually. The latter refers to a window which requires user interaction
to bring forward.


\paragraph{Searchable Toolspace} This is a boolean variable which indicates
whether the total set of available tools, components, or actions offered by
the \ac{ide} can be searched through by name or keyword.


\paragraph{Toolbar Styles} The toolbar styles of an \ac{ide} refer to the set of
user interface component idioms employed by the \ac{ide}.~\cite{galitz2007} This
nominal variable can take combinations of multiple values. Example values
might include \texttt{Icons}, \texttt{Menus}, \texttt{Ribbons},
\texttt{Trees}.


\paragraph{Visual Clutter} The clutter of an interface is defined as the
number and organization of tools available on the screen versus the amount
of workspace provided by the \ac{ide}. If the \ac{ide} offers no method for tool
organization or if there is an immense amount of tools visible at once,
then the \ac{ide} is likely to be visually cluttered. Visual clutter is a
nominal variable and can take the values \texttt{Low}, \texttt{Medium}, or
\texttt{High}. \Sect{subsec:ideevaluation} includes details on the proper
evaluation of this qualitative feature.


\subsubsection{Human Interface Features} \label{subsubsec:humaninterface}

The human interface features of an \ac{ide} include aspects of the software
interface that affect how the human user interacts with the \ac{ide}, either
mechanically (\eg through physical devices and media) or mentally (\eg the
mental load required of the user to operate the \ac{ide}).


\paragraph{Essential Efficiency} The essential efficiency of an \ac{ide}
measures the level to which the system automates tasks for the user. It is
calculated as the ratio of the number of steps in a concrete use case to
the number of steps in the representative use case (see
\Eq{eq:eefficiency}). This metric, unlike those for automation put forth in
\cite{wei1998}, does not require user experience reports and can be
measured through simple use case analysis.
%
\begin{align}\label{eq:eefficiency}
  1 - \frac{\text{Number of steps in concrete use case}}
            {\text{Number of steps in essential use case}}
\end{align}


\paragraph{Interface Efficiency} The interface efficiency of an \ac{ide} is a
concept related to the productivity of an interface. It measures the number
of physical actions (including keystrokes, mouse clicks, and fine mouse
movements) required of the operator to complete a task compared against
the number of abstract steps in the representative essential use case.

This metric is different from the Essential Efficiency proposed in
\cite{constantine1996} because it studies physical user actions instead of
concrete task steps. Also note that this variable can take negative values,
indicating that the number of physical actions required to complete an
essential use case exceeds the number of abstract steps in the use case.
%
\begin{align}\label{eq:iefficiency}
  1 - \frac{\text{Number of physical actions to complete use case}}
            {\text{Number of steps in essential use case}}
\end{align}


\paragraph{Keyboard Use} Keyboard Use refers to the extent to which an \ac{ide}
utilizes the use of a keyboard. This can range from a complete absence of
any keyboard actions to providing certain actions which only a keyboard can
perform. This nominal variable can take one of four values. A value of
\texttt{None} indicates that no keyboard use is supported. \texttt{Simple}
indicates that the keyboard is used only for typing annotations,
properties, or comments. \texttt{Optional} indicates that the option to use
the keyboard to execute some actions is present, but these actions can also
be completed using a mouse. \texttt{Required} indicates that there are
actions which can only be completed through the use of the keyboard, no
mouse equivalent is available.


\paragraph{Mode of Element Creation} This describes the process through
which the user creates elements in the \ac{ide}. The \emph{Drag n Drop} process
refers to a single mouse press event followed by a dragging motion of the
mouse and completed when the mouse button is released. The \emph{Point n
Click} process utilizes a single mouse click to indicate a selection and
followed with subsequent mouse clicks elsewhere to define placement.

This nominal variable can take one of four possible values created by
combining \texttt{Drag n Drop} or \texttt{Point n Click} with the
multiplicity of the action: either \texttt{(1:1)} or \texttt{(1:n)}. The
former multiplicity indicates that a single element is created for each
action, while the latter indicates that multiple can be created after for
action.


\paragraph{Tertiary Interface Devices} This nominal variable describes any
third party human interface devices which can be used to interact with the
\ac{ide}. This could include audio devices such as MIDI keyboards or
microphones, mobile integration, etc. Variable values are the type of
tertiary devices allowed for the \ac{ide}.


\subsubsection{Integration Features} \label{subsubsec:integration}

Integration is the manner with which the \ac{ide} integrates with the visual
language it supports. This includes any visual representation of language
syntax or semantics, as well as any tools to assist the user with
understanding the supported language.


\paragraph{Allowed Relations Indicated} This refers to an \ac{ide}'s ability to
emphasize possible syntactically correct connection points. This is often
demonstrated with either the highlighting of allowed relations or the
dimming of impossible relations. Boolean values indicate whether the \ac{ide}
supports this feature or not.


\paragraph{Output Generation Style} This nominal variable describes the
mode with which the \ac{ide} renders and displays output to the user. It is a
dual axis nominal variable, measuring whether output is direct or indirect
as well as live or caused by a trigger. \texttt{Direct/indirect} describes
whether the user directly modifies output or acts via a layer of
abstraction (\eg via a model). \texttt{Live/trigger} describes whether
output is generated and displayed live or after some event triggered by the
user.


\paragraph{Syntax Enforcement} The level of syntax enforcement of an \ac{ide}
describes the mode with which the \ac{ide} enforces its supported language's
syntax requirements, if at all. The nominal value \texttt{Explicit}
indicates that the \ac{ide} explicitly enforces syntax requirements by
indicating to the user the presence of any syntax errors. \texttt{Implicit}
enforcement indicates that the \ac{ide} does not allow syntactically illegal
operations to occur in the first place by means of some structural
mechanism. The value \texttt{None} indicates that the \ac{ide} does not support
syntax enforcement and the user is required to review their models' syntax
manually.


\subsubsection{Language Syntax Features} \label{subsubsec:languagesyntax}

Language syntax variables describe properties of the supported visual
language syntax. While not strictly components of the \ac{ide}, they are
intimately tied to the overall style of the \ac{ide} and thus included in this
study.


\paragraph{Complexity Management} Any characteristics or features of the
visual language that serve to reduce the complexity of that language.
Reducing complexity refers specifically to decreasing the level of
\emph{diagrammatic complexity} while maintaining information transfer to
the user.~\cite{moody2009} This nominal variable can take one of three
possible values. \texttt{Modularization} indicates that large systems
within the language are divided into smaller subsystems to reduce
complexity.~\cite{moody2009} \texttt{Hierarchy} indicates that systems
within the language can be represented at different levels of detail.
\cite{moody2009} \texttt{None} indicates that the visual language does not
support complexity management functionality.


\paragraph{Connection Style} A language's connection style refers to the
manner with which connections between elements are displayed. This dual
axis nominal variable measures connections as \texttt{overlapping}
\emph{vs.} \texttt{linked} as well as connection sources as \texttt{point}
\emph{vs.} \texttt{region} based.~\cite{costagliola2002} The former axis
describes the visual representation of the connections, while the latter
describes how links are connected. If the supported language does not fall
on either of these axes, it is not connection based~\cite{costagliola2002}
and this feature takes the value \texttt{Geometric}.


\paragraph{Degree of Language Visual Richness} The degree of a language's
visual richness describes the extent to which it utilizes visual
variables to increase the visual discriminability of elements. This
is a compound variable composed of ten sub-variables. Each sub-variable
is a boolean measure determining whether or not the described visual
variable is utilized in the interface to distinguish available tools.

\subparagraph{Icons} Images contained within elements are used to represent
different elements.~\cite{costagliola2002,moody2009}

\subparagraph{Shape} Distinct shapes indicate different elements.
\cite{moody2009}

\subparagraph{Size} Different element sizes indicate distinct elements.
\cite{moody2009}

\subparagraph{Color} Color is used to indicate distinct elements.
\cite{moody2009}

\subparagraph{Text} Text (or typographic variation) is used to identify or
distinguish elements.~\cite{moody2009}

\subparagraph{Texture} Shading or shadows are used to modify elements or to
distinguish between distinct elements.~\cite{moody2009}

\subparagraph{Brightness} The brightness of a color (\ie its perceived
luminosity) is used to indicate a difference between elements.
\cite{moody2009}

\subparagraph{Orientation} Rotation of an element indicates differences
between elements.~\cite{moody2009}

\subparagraph{Horizontal Position} The horizontal location of an element
indicates differences between elements.~\cite{moody2009}

\subparagraph{Vertical Position} The vertical location of an element
indicates differences between elements.~\cite{moody2009}


\subsection{\acs{ide} Evaluation} \label{subsec:ideevaluation}

With \ac{ide} feature definitions established, each \ac{ide} was evaluated
for its performance according to each feature. The vast majority of
features resulted in simple classification or binary evaluation tasks
depending on the characteristics of the \acp{ide}. However, three features
required more intensive evaluation. The processes used for these features
are described in \Sect{subsubsec:mturk} and \Sect{subsubsec:efficiency}.

\subsubsection{\acl{mt} User Study} \label{subsubsec:mturk} To evaluate the
visual clutter of an \ac{ide} requires qualitative user feedback.
Amazon.com's \ac{mt} was used as a crowd-sourcing platform to perform the
user study and gauge opinions of visual clutter. We generated three unique
screenshots of each \ac{ide} and darkened the workspace area to remove
attention from diagrammatic complexity. These images were distributed
as \ac{mt} \acp{hit} requiring five unique raters per \ac{hit}. On average,
each worker spent about one minute studying and rating an image on a scale
of one (low clutter) to five (high clutter). Each \ac{hit} rewarded its
worker with \$0.02. \ac{ide} rating values were retrieved and averaged
between each reviewer for one screenshot and then each screenshot for one
\ac{ide}. Upon completion, the \acp{hit} were rated, in total, by 12
anonymous workers.

Because multiple reviewers rated each image, we performed an \ac{irr}
measure to ensure agreement across reviewers. Using the R statistical
library \texttt{irr} we perform a two-way, agreement, average-measure
\ac{icc}. The result, \ac{icc} = 0.648, is within the ``good'' range of
significance.~\cite{cicchetti1994,hallgren2012} This \ac{icc} value
indicates that the reviewers were, in general, in agreement about their
ranking of interface clutter. Note that there were more than five
participating reviewers total, despite five sets of reviews (\ie the
experimental design is not ``fully crossed''~\cite{hallgren2012}). We argue
that this is not significant, however, because the five sets of reviewers
are disjoint sets, acting as entirely independent actors.

\subsubsection{Efficiency Study} \label{subsubsec:efficiency} The
evaluation of both interface and essential efficiency involved the creation
of an abstract use case for each \ac{ide} and subsequent evaluation of the
corresponding concrete use case. Three abstract use cases were developed
for each \ac{ide} in order of increasing task complexity. The first two use
cases were common between all \acp{ide} while the third use case differed
slightly. Each abstract use case was then manually performed within the
\acp{ide} to generate corresponding concrete use cases. Every physical
action required for completion was tracked (specifically keystrokes, mouse
clicks, and fine mouse movements).

After all concrete use cases were completed, it was determined that the
third (most complex) use case was the most representative use case for each
\ac{ide} because concrete use case steps did not vary greatly for the least
complex use cases. The number of steps in the concrete use cases (\ie the
number of distinct tasks the \ac{ide} required the user to keep in mind and
perform) became the numerator for the essential efficiency ratio. The
number of physical actions performed became the numerator for the interface
efficiency ratio.
