\section{Results}
\label{sec:results}

Tables~\ref{tab:chrome}--\ref{tab:integration-langsyn} outline the results of the evaluation of the 25 \acp{ide} we considered in this study.
We give further explanations below for each in alphabetical order.
%
%\begin{landscape}
\begin{table*}
	\centering
		{\scriptsize
		\begin{tabular}{l|cccccccc}
		  \hline\hline
		  &&&&&&&&		\\[-1.5ex]
      \textbf{IDE}
      & \textbf{\scshape Features}
      & \textbf{\scshape Context}
      & \textbf{\scshape ToolRichness}
      & \textbf{\scshape Perspectives}
      & \textbf{\scshape Properties}
      & \textbf{\scshape Searchable}
      & \textbf{\scshape ToolStyle}
      & \textbf{\scshape Clutter}			\\[1ex]
		  \hline
		  &&&&&&&&		\\[-1.5ex]
      \textbf{Alice3} & 8	& Yes	& 6	& 2	& Omnipresent	& No	& Tabs	& 3.40			\\[1ex]
      \textbf{AToMPM}    & 4	& No	& 5	& 1	& Manual	& No	& Modular	& 2.60			\\[1ex]
      \textbf{AudioMulch}    & 7	& Yes	& 6	& 2	& Manual	& No	& Tree, Windows	& 3.13			\\[1ex]
      \textbf{Blender}    & 7	& Yes	& 7	& Multiple	& Omnipresent	& No	& Multiple	& 4.00			\\[1ex]
      \textbf{\cameleon}    & 3	& Yes	& 5	& 1	& Manual	& Yes	& Tree	& 2.07			\\[1ex]
      \textbf{EMF}    & 9	& Yes	& 5	& Multiple	& Manual	& Yes	& Tree, Icons	& 4.07			\\[1ex]
      \textbf{GNU Radio Companion}    & 6	& No	& 4	& 1	& Manual	& Yes	& Icons	& 2.80			\\[1ex]
      \textbf{Grasshopper 3D}    & 7	& Yes	& 4	& 1	& None	& Yes	& Ribbons	& 2.80			\\[1ex]
      \textbf{Max}    & 7	& No	& 4	& 2	& Manual	& Yes	& Tree, Windows	& 2.73			\\[1ex]
      \textbf{MetaEdit+}    & 7	& Yes	& 4	& 3	& Manual	& No	& Icons, Windows	& 1.93			\\[1ex]
      \textbf{MIT AppInventor2}    & 3	& Yes	& 4	& 2	& Omnipresent	& No	& Drawers	& 3.27			\\[1ex]
      \textbf{MST}    & 6	& No	& 3	& 1	& Manual	& No	& Icons	& 2.40			\\[1ex]
      \textbf{Piet Creator}    & 4	& No	& 7	& 1	& None	& No	& Icons	& 2.13			\\[1ex]
      \textbf{Scratch}    & 3	& No	& 6	& 2	& None	& No	& Drawers	& 3.53			\\[1ex]
      \textbf{SimuLink}    & 7	& No	& 3	& 1	& Manual	& Yes	& Icons, Tree	& 3.80			\\[1ex]
      \textbf{Stencyl}    & 6	& Yes	& 8	& Multiple	& Manual	& Yes	& Tabs, Icons	& 2.73			\\[1ex]
      \textbf{Tersus}    & 8	& No	& 3	& 1	& Omnipresent	& No	& Drawers	& 3.20			\\[1ex]
      \textbf{TouchDevelop}    & 6	& Yes	& 6	& 1	& None	& No	& Icons	& 3.60			\\[1ex]
      \textbf{UMLet}    & 6	& No	& 3	& 1	& Omnipresent	& No	& Palette	& 3.00			\\[1ex]
      \textbf{VioletUML}    & 6	& No	& 4	& 1	& Manual	& No	& Drawers	& 2.47			\\[1ex]
      \textbf{VisSim}    & 8	& No	& 4	& 1	& Manual	& Yes	& Tree, Menu	& 3.33			\\[1ex]
      \textbf{Visual Paradigm}    & 7	& Yes	& 5	& 1	& Omnipresent	& No	& Icons, Menu	& 3.67			\\[1ex]
      \textbf{Visual Use Case}    & 3	& No	& 4	& 7	& Manual	& No	& Icons, Tree	& 2.67			\\[1ex]
      \textbf{WebRatio}    & 8	& No	& 4	& 3	& Omnipresent	& No	& Icons, Menu	& 3.93			\\[1ex]
      \textbf{YAWL}    & 5	& No	& 5	& 2	& Omnipresent	& No	& Icons	& 2.27			\\[1ex]
			%\hline
		\end{tabular}}
	\caption{Measure of the chrome of \acp{ide}.}
	\label{tab:chrome}
\end{table*}
%
\begin{table*}
	\centering
		{\scriptsize
		\begin{tabular}{l|cc|ccccc}
		  \hline\hline
		  &&&&&&&		\\[-1.5ex]
      \multirow{2}{*}{\textbf{IDE}}
      & \multicolumn{2}{c|}{\textbf{\scshape Audience}}
      & \multicolumn{5}{c}{\textbf{Human Interface}}  \\[1ex]
      & \textbf{\scshape Domain}
      & \textbf{\scshape Skill}
      & \textbf{\scshape EEfficiency}
      & \textbf{\scshape IEfficiency}
      & \textbf{\scshape Keyboard}
      & \textbf{\scshape Mode}
      & \textbf{\scshape Devices}			\\[1ex]
		  \hline
		  &&&&&&&		\\[-1.5ex]
      \textbf{Alice3}    & Animation	& Novice	& 0.46	& 0.08	& Optional	& Drag n Drop (1:1)	& None			\\[1ex]
      \textbf{AToMPM}    & Modeling	& Intermediate	& 0.23	& 0.15	& Optional	& Point n Click (1:n)	& None			\\[1ex]
      \textbf{AudioMulch}    & Music	& Expert	& 0.31	& 0.31	& Simple	& Drag n Drop (1:1)	& Keyboards			\\[1ex]
      \textbf{Blender}    & 3D Modeling	& Expert	& 0	& -0.2	& Required	& Point n Click (1:1)	& None			\\[1ex]
      \textbf{\cameleon}    & Prototyping	& Intermediate	& 0.23	& 0.31	& Required	& Drag n Drop (1:1)	& None			\\[1ex]
      \textbf{EMF}    & Software	& Expert	& 0.23	& 0.62	& Optional	& Point n Click (1:1)	& None			\\[1ex]
      \textbf{GNU Radio Companion}    & Software	& Intermediate	& 0.38	& 0.62	& Optional	& Drag n Drop (1:1)	& None			\\[1ex]
      \textbf{Grasshopper 3D}    & 3D Modeling	& Intermediate	& 0.31	& 0.31	& Optional	& Point n Click (1:1)	& None			\\[1ex]
      \textbf{Max}    & Music	& Expert	& 0.15	& 0.15	& Simple	& Drag n Drop (1:1)	& Microphones			\\[1ex]
      \textbf{MetaEdit+}    & Modeling	& Intermediate	& 0.31	& -0.08	& Simple	& Point n Click (1:1)	& None			\\[1ex]
      \textbf{MIT AppInventor2}    & Software	& Novice	& 0.46	& 0.54	& Simple	& Drag n Drop (1:1)	& Mobile			\\[1ex]
      \textbf{MST}    & Simulation	& Intermediate	& 0.38	& 0.38	& Simple	& Point n Click (1:1)	& None			\\[1ex]
      \textbf{Piet Creator}    & Software	& General	& 0.17	& -0.33	& None	& Point n Click (1:n)	& None			\\[1ex]
      \textbf{Scratch}    & Software	& Novice	& 0.46	& 0.54	& Simple	& Drag n Drop (1:1)	& None			\\[1ex]
      \textbf{SimuLink}    & Simulation	& Expert	& 0.31	& 0.15	& Simple	& Drag n Drop (1:1)	& None			\\[1ex]
      \textbf{Stencyl}    & Software	& Novice	& 0.46	& 0.54	& Simple	& Drag n Drop (1:1)	& Mobile			\\[1ex]
      \textbf{Tersus}    & Software	& Intermediate	& 0.31	& 0.15	& Simple	& Point n Click (1:1)	& None			\\[1ex]
      \textbf{TouchDevelop}    & Software	& Novice	& 0.44	& 0.33	& None	& Point n Click (1:1)	& Mobile			\\[1ex]
      \textbf{UMLet}    & Modeling	& General	& 0.31	& -0.15	& Simple	& Drag n Drop (1:1)	& None			\\[1ex]
      \textbf{VioletUML}    & Modeling	& Intermediate	& 0.15	& 0.15	& Simple	& Point n Click (1:n)	& None			\\[1ex]
      \textbf{VisSim}    & Simulation	& Expert	& 0.31	& 0.31	& Simple	& Drag n Drop (1:1)	& None			\\[1ex]
      \textbf{Visual Paradigm}    & Modeling	& Intermediate	& 0.23	& 0.23	& Simple	& Drag n Drop (1:1)	& None			\\[1ex]
      \textbf{Visual Use Case}    & Modeling	& Intermediate	& 0.54	& 0.38	& Simple	& Point n Click (1:1)	& None			\\[1ex]
      \textbf{WebRatio}    & Software	& Intermediate	& 0.23	& 0.23	& Simple	& Point n Click (1:1)	& None			\\[1ex]
      \textbf{YAWL}    & Workflow	& Expert	& 0.31	& 0.31	& Simple	& Point n Click (1:n)	& None			\\[1ex]
			%\hline
		\end{tabular}}
	\caption{Measure of the human interface of \acp{ide} and their intended audience.}
	\label{tab:interface}
\end{table*}
%
\begin{table*}
	\centering
		{\scriptsize
		\begin{tabular}{l|ccc|ccc}
		  \hline\hline
		  &&&&&&		\\[-1.5ex]
      \multirow{2}{*}{\textbf{IDE}}
      & \multicolumn{3}{c|}{\textbf{Integration}}
      & \multicolumn{3}{c}{\textbf{Language Syntax}}  \\[1ex]
      & \textbf{\scshape Relations}
      & \textbf{\scshape Output}
      & \textbf{\scshape Syntax}
      & \textbf{\scshape Complexity}
      & \textbf{\scshape Connection}
      & \textbf{\scshape Language Richness}			\\[1ex]
		  \hline
		  &&&&&&		\\[-1.5ex]
      \textbf{Alice3}    & No	& Direct Triggered	& Implicit	& Modularization	& Overlapping Regions	& 8			\\[1ex]
      \textbf{AToMPM}    & No	& Indirect Triggered	& Explicit	& None	& Linked Regions	& 6			\\[1ex]
      \textbf{AudioMulch}    & Yes	& Indirect Live	& Implicit	& None	& Linked Points	& 3			\\[1ex]
      \textbf{Blender}    & No	& Direct Live	& Implicit	& Modularization	& Geometric	& 9			\\[1ex]
      \textbf{\cameleon}    & Yes	& Direct Live	& Explicit	& Hierarchy	& Linked Points	& 5			\\[1ex]
      \textbf{EMF}    & No	& Indirect Triggered	& Explicit	& Modularization	& Linked Points, Linked Regions	& 5			\\[1ex]
      \textbf{GNU Radio Companion}    & No	& Indirect Triggered	& Explicit	& None	& Linked Points	& 2			\\[1ex]
      \textbf{Grasshopper 3D}    & No	& Indirect Live	& Explicit	& None	& Linked Points	& 6			\\[1ex]
      \textbf{Max}    & No	& Indirect Live	& None	& Modularization	& Linked Points	& 5			\\[1ex]
      \textbf{MetaEdit+}    & No	& Direct Live	& Implicit	& Modularization	& Linked Regions	& 5			\\[1ex]
      \textbf{MIT AppInventor2}    & No	& Direct Triggered	& Implicit	& Modularization	& Overlapping Regions	& 9			\\[1ex]
      \textbf{MST}    & No	& Direct Live	& None	& None	& Overlapping Points	& 4			\\[1ex]
      \textbf{Piet Creator}    & No	& Indirect Triggered	& None	& None	& Geometric	& 4			\\[1ex]
      \textbf{Scratch}    & No	& Direct Triggered	& Implicit	& Modularization	& Overlapping Regions	& 10			\\[1ex]
      \textbf{SimuLink}    & No	& Indirect Live	& None	& None	& Linked Regions	& 4			\\[1ex]
      \textbf{Stencyl}    & No	& Direct Triggered	& Implicit	& Modularization	& Overlapping Regions	& 10			\\[1ex]
      \textbf{Tersus}    & No	& Direct Triggered	& None	& Hierarchy	& Linked Points, Linked Regions, Geometric	& 3			\\[1ex]
      \textbf{TouchDevelop}    & No	& Indirect Triggered	& Explicit	& Modularization	& Geometric	& 4			\\[1ex]
      \textbf{UMLet}    & No	& Direct Live	& None	& None	& Overlapping Points	& 4			\\[1ex]
      \textbf{VioletUML}    & No	& Direct Live	& Explicit	& None	& Linked Regions	& 4			\\[1ex]
      \textbf{VisSim}    & No	& Direct Live	& None	& Hierarchy	& Linked Points	& 4			\\[1ex]
      \textbf{Visual Paradigm}    & No	& Direct Live	& Implicit	& None	& Linked Regions	& 2			\\[1ex]
      \textbf{Visual Use Case}    & No	& Direct Live	& None	& Modularization	& Linked Regions	& 3			\\[1ex]
      \textbf{WebRatio}    & No	& Indirect Triggered	& None	& Hierarchy	& Linked Regions, Geometric	& 4			\\[1ex]
      \textbf{YAWL}    & No	& Direct Triggered	& None	& None	& Linked Regions	& 4			\\[1ex]
			%\hline
		\end{tabular}}
	\caption{Measure of the integration and language syntax of \acp{ide}.}
	\label{tab:integration-langsyn}
\end{table*}
%
\paragraph{Alice3} Designed primarily as an educational programming
environment, Alice3 alienates skilled users by expecting a lower level of
skill in its domain. It makes up for this loss in accessibility, however,
with its wide, almost universal, support of common \ac{ide} features as
well as context-sensitive tooling. Alice3 provides a medium level of visual
richness in its interface chrome, but boasts one of the highest essential
efficiency values. This efficiency value is achieved through the ``MIT
Scratch'' style of visual syntax. However, despite greatly reducing
operator mental load, its visually cluttered design only manages a
near-neutral interface efficiency. The introduction of optional keybindings
is a redeeming factor for Alice3 and, like most educational programming
\acp{ide}, Alice3 includes simple modular complexity management and a
visually appealing level of language visual richness.

\paragraph{AToMPM} AToMPM is unique among the \acp{ide} in this study in
that it is built as a modeling framework specifically to develop domain-specific modeling \acp{ide}.
While AToMPM does not support a large number of popular \ac{ide} features
or visual richness variables, its modular toolbar system helps to reduce
visual clutter by loading only the features the user deems necessary.
AToMPM allows for a small number of optional keybindings which can help
increase utility, but essential efficiency and interface efficiency are
only slightly above a one-to-one ratio with the essential use cases.
Explicit syntax enforcement is provided which protects the user from
potential illegal operations, but no complexity management system is in
place to handle user mental load. Because AToMPM is used to develop
\acp{ide} it can be argued that resulting generated AToMPM tools may score
differently, but overall AToMPM achieves roughly average performance
compared to other \acp{ide} in the study.

\paragraph{AudioMulch} Although AudioMulch offers a wide array of tools and
an in-depth interface ideal for professionals in the music industry, the
overall complexity of the design greatly reduces the accessibility for
anyone else. Nevertheless, it supports a large amount of popular \ac{ide}
features and offers equally high essential and interface efficiency
ratings. This is visually assisted by a relation-highlighting feature which
also provides AudioMulch with an implicit syntax enforcement, both of which
greatly aid the user in model creation. Unfortunately, there is no effort
made to manage the high amount of complexity within the \ac{ide} and
virtually all of the canvas elements look exactly the same, ultimately
awarding AudioMulch with a low language visual richness score.

\paragraph{Blender} Designed for a skilled target audience of experts in
the domain, it is no surprise that Blender supports most common \ac{ide}
features as well as context-sensitive tooling. Its high level of chrome
visual richness and the large number of perspectives available relative to
the average found in this study also contribute to a high quality
interface. However, the fact that it offers so many features leads directly to the second highest
observed value for visual clutter. Blender possesses no particular efficiency
techniques, remaining around a perfect one-to-one relationship with the
measured essential use cases. Its heavy use of the keyboard reduces
accessibility to a wider audience, although the target skill level is
already a limiting factor. Finally, Blender provides modularization
complexity management through saving and duplication tools, along with one
of the most visually rich languages observed.

\paragraph{\cameleon} Intended for rapid, flexible visual prototyping of
functional algorithms, \cameleon~is relatively feature impoverished --- it
only supports three of the most popular \ac{ide} features. It does however
boast a large number of available tools that are conveniently searchable.
The interface chrome of \cameleon~employs five visual richness variables,
slightly above the average count. Its low visual clutter interface provides
a positive level of essential and interface efficiency. Interaction
requires use of the keyboard for non-essential actions, specifically
advanced navigation and zooming. \cameleon~supports the user by
interactively highlighting allowed syntactic relations and explicitly
enforcing syntax requirements. More support is provided by its hierarchical
complexity management system. Overall, \cameleon~is an efficient, simple
tool with a large library of functionality and a focus on supporting syntax
requirements.

\paragraph{\acl{emf}} EMF is one of the most sophisticated
\acp{ide} in this study and demands an expert level of skill from target
users. It is also the only \ac{ide} studied which supports all of the
measured popular features, as any expert in the domain would likely come to
expect. The \ac{emf} interface chrome supports five visual richness
variables as well as a large number of predefined interface perspectives.
The high level of supported perspectives, while perhaps intimidating,
increases the overall power and utility of the interface. Additionally, the
\ac{emf} tool space is searchable, which counterbalances the highest
observed level of visual clutter. Despite being extremely cluttered,
\ac{emf} has average levels of essential efficiency and the highest
measured level of interface efficiency. The supported visual languages do
not make use of more than five visual variables but do allow for complexity
management. Overall, \ac{emf} is well suited for an expert user with high
efficiency and several convenience features, but is too cluttered and not
visually rich enough to support a wider audience of varied skill levels.

\paragraph{\acl{grc}} \ac{grc} is a simple platform designed to aid the
development of signal processing software without the need to understand or
write code. It does, however, require a moderate level of skill in the
domain. \ac{grc} supports slightly more than the average number of popular
\ac{ide} features as well as the ability to search through available tools.
\ac{grc} is not extensively visually pleasing, however, as the interface
chrome only utilizes four visual variables while the supported language
only employs two. Despite this, the interface is highly efficient, with the
highest observed interface efficiency and a correspondingly high level of
essential efficiency. It also supports optional keyboard use and
explicit syntax checking. \ac{grc} is designed as an easy-to-use interface
for non-programmers and manages to maintain simplicity while still offering a
large amount of technical power.

\paragraph{Grasshopper 3D} Though Grasshopper is able to provide the user
with a relatively simple and easy-to-use interface, beginners would likely
shy away from the complexity of its core functionality. Even so, it offers
a high amount of popular \ac{ide} features and context sensitive tools, as
well as the ability to search through its vast library of tools quickly and
easily by name. Grasshopper also possesses very good efficiency techniques,
maintaining both values at a more-than-decent level. In addition, the
optional use of a keyboard is supported, offering functionality on another
level to increase accessibility. However, complexity management is not supported at
all and visual richness in both the language and tools are
mediocre at best.

\paragraph{Max} Its overly simple appearance can be misleading---it is
very much designed for skilled users in the music development domain. Max
supports several popular features and an average number of visual
richness variables in both its chrome and supported language, making it
accessible and relatively visually pleasing. Its use of a searchable
toolspace is unique among the studied \acp{ide}: in order to create
almost any advanced element in the workspace, the user must search by name
or description for the element. Only a small subset of its functionality
can be reached otherwise. This simultaneously radically reduces visual
clutter by hiding most elements from the user, while also drastically
increasing mental overhead for users. Max offers a below average level of
interface and essential efficiency, as well as a convenient modular
complexity management system. Once mastered, Max can be a powerful and
efficient tool for developing music and audio processing tools, but the
amount to which the library of functionality is removed from the user can
be intimidating and, ultimately, greatly decreases accessibility.

\paragraph{MetaEdit+} As for AToMPM, MetaEdit+ is a tool for creating domain-specific modeling environments.
Aimed towards an intermediate level of users, the
many different tasks and steps that go into the design of a simple model in
MetaEdit+ can easily be daunting for newer users.  These tasks are
fortunately divided through modularization, allowing them to be much more
manageable within the \ac{ide}. Furthermore, many of the popular \ac{ide}
features as well as context sensitive tools are present in the interface,
increasing the accessability even more. MetaEdit+ also holds the lowest
clutter value and integrates an inplicit syntax enforcement, providing the
user with a clear and easy-to-use workspace. A favorable essential
efficiency value is also present, whereas the interface efficiency suffers
from its necessity to complete a dialog for each created element.

\paragraph{MIT AppInventor} Much like Alice3 and the other educational
interfaces, MIT's AppInventor is designed for an unskilled, novice
audience, reducing the breadth of target audience. By only supporting three
of the top ten popular features, AppInventor additionally alienates its
audience through nonconformity with expected standards. While colorful,
AppInventor's interface only supports four visual variables to distinguish
elements in the chrome, relying heavily on icons and text. It
combines the ``MIT Scratch'' style of visual syntax along with a Drag n
Drop element creation workspace, resulting in high levels of effective and
interface efficiency. Another artifact of the ``MIT Scratch'' style of
visual syntax is its implicit syntax enforcement, assisting the user by
preventing illegal structures. Finally, the high level of language visual
richness and modular complexity management scheme result in an overall
visually pleasant experience.

\paragraph{MST Workshop} Though MST Workshop offers a very simple and easy
to use interface, it does not offer much explanation as to the use of its
many different simulation categories. As such, they are virtually unusable
by anyone without prior knowledge of that subject, drastically limiting its
accessibility. In the same vein, it does not incorporate enough visual
richness variables to easily discern or interpret the tools and very few
popular \ac{ide} features are supported. Though the efficiency values are
above average and the clutter was rated to be relatively low, no
actions were made to reduce the complexity of the system or even enforce
the language's syntax. All of this merged together with a less-than-stellar
language visual richness level shows that MST Workshop definitely has room
for improvement.

\paragraph{Piet Creator} As the primary \ac{ide} used in the creation of
Piet programs, Piet Creator provides a very simple
interface for novices without limiting the usability for more expert users.
On top of that, the toolbars utilize seven of the eight tool visual
richness variables and it possesses an extremely low clutter rating,
maximizing accessibility for all users.  However, Piet Creator does not
offer any sort of properties dialog, eliminating the ability to manage data
on a deeper level. The user is also limited to using only the mouse for
every task, which creates a poor combination with the fact that Piet
Creator holds the lowest interface efficiency value. The Piet language
syntax is not enforced in the slightest, forcing the user to manually debug
his whole program to locate any error in the code. In general, the
simplicity of the design allows for a visually
appealing environment, but creates a dearth of features.

\paragraph{Scratch} One of the earliest educational interfaces present in
this study, Scratch has influenced a great deal of later \acp{ide},
including Stencyl, AppInventor, and Alice3 to name a few. Scratch is
specifically targeted toward a novice audience with a low skill level, and
only supports three of the top ten popular features. The interface employs
an average level of visual richness variables, and does not even define an
object properties window. The sparse, cluttered interface redeems itself
through high values of essential and interface efficiency. Implicit syntax
enforcement provides a safe environment for learning users, and the highest
observed level of language visual richness provides an engaging, visually
rich display. The focus on efficiency and visual richness works well in an
\ac{ide} designed for education by accelerating reinforcement and engaging
student attention.

\paragraph{Simulink} Though Simulink features a plethora of components that
can allow it to perform virtually any electrical simulation, the vast scope
of its functions and the amount of on-screen tasks greatly reduces its
accessibility. Each individual function that Simulink provides creates its
own dialog window on screen, severely increasing visual clutter and
complexity with prolonged use and no techniques for complexity management.
A searchable toolspace is present to take off some mental load in dealing
with Simulink's huge tool libraries, however very few visual richness
variables are integrated to increase discernibility between the tools. The
canvas elements also provide some relatively good efficiency values, though
the \ac{ide} provides no features to enforce the syntax of the simulation
language. The lack of visual richness overall detracts from the usability
and enjoyment of this tool, despite it being very powerful.

\paragraph{Stencyl} It provides a very easy-to-use interface for game
software creation for an early level audience, but in the process
sacrifices any higher level functionality. This \ac{ide} also integrates
the ``MIT Scratch'' style of coding, which carries along with it very high
efficiency values and implicit syntax enforcement. Every tool visual
richness variable is utilized within the toolbars, which are also
searchable, allowing for a very user-friendly and productive interface.
Each language visual richness variable is supported within the canvas as
well, thereby giving Stencyl perfect visual richness ratings. A very large
number of perspectives are present due to Stencyl's powerful modularization
techniques, greatly reducing the amount of mental strain on the user to
handle the many steps that go into creating a game and showing Stencyl to
be an altogether organized and well-designed \ac{ide}.

\paragraph{Tersus} It is able to offer an interface for web application
design with a very large number of popular \ac{ide} features, largely
thanks to the integration of Eclipse's user interface. Unfortunately, very
few visual richness variables are utilized for either the tools or the
language and there is no syntax enforcement to aid the user at all through
the design process. The efficiency of the \ac{ide} is at an acceptable level.
Tersus hits a high point with its use of a hierarchical design to program
behaviors and functions of the various components in the model. Though this
is a very useful feature, it does not distract at all from the overall
blandness of the language.

\paragraph{TouchDevelop} Designed as an educational program as part of
Microsoft's ``Hour of Code" campaign, TouchDevelop offers a simple user
interface with a very limited functionality. Though the model itself is
textually code based, the code is written by pressing buttons representing
different commands, objects, attributes, etc., increasing both essential
and interface efficiency by a good margin. However, the large, on-screen keyboard
gives the interface a slightly cluttered feel and the text-based
model places limits on the amount of language visual richness variables
that can be integrated in the design.

\paragraph{UMLet} UMLet provides a general purpose, easy-to-use software
interface with no particular expectations about user skill level in \ac{uml}. It offers
many of the most popular \ac{ide} features a user might expect, but does
not offer a visually rich chrome or language. Conceptually efficient but
ultimately visually cluttered, UMLet measured high in essential efficiency but
low in interface efficiency. To the detriment of the inexperienced user,
UMLet offers no syntax enforcement. It also offers no complexity management
devices to ease mental load. Barebones to the extreme, UMLet provides users
with a visually simple interface and no additional usability features like
context-sensitive tooling or searchable tools. Its power, however, lies in
its simplicity and ease of use at any entry skill level.

\paragraph{Violet} Similar to UMLet, Violet provides an easy-to-use
interface with no specific entry skill level assumptions about \ac{uml}. It offers the
same common \ac{ide} features as UMLet, with a slightly richer chrome.
The interface of Violet is slightly less cluttered than in UMLet, but also offers
a lower essential efficiency. Violet is higher in interface efficiency over
UMLet, though. Neither interface offers syntax enforcement or complexity
management and, since both support \ac{uml} as the target visual language,
both share a low level of language visual richness. The two interfaces are
relatively featureless and differ primarily in the mode of user
interaction.

\paragraph{VisSim} VisSim is a highly sophisticated visual simulation tool,
designed for experts in the field. By supporting most popular \ac{ide}
features, VisSim meets most developer expectations. The interface and
language are adequately visually rich, supporting the average count of four
visual variables. The large library of functionality is conveniently
searchable, and the interface is highly efficient overall. VisSim has an easy-to-use
hierarchical complexity management system which reduces overall mental
load, but a lack of syntax enforcement may leave users in a dangerous
position, forcing them to rely on verification at run-time only.

\paragraph{Visual Paradigm} It offers a \ac{uml}-based
interface designed to support every defined \ac{uml} diagram. As such,
however, much of the functionality may be daunting to a novice to \ac{uml}
architecture. Many popular \ac{ide} features are present, though, as well
as context sensitive tools for a slightly more accessible use. The design of Visual
Paradigm allows for implicit syntax enforcement, a very useful
feature when dealing with \ac{uml}, with an acceptable interface
efficiency and average essential efficiency value. An unfavorable amount of
clutter is present in the \ac{ide} however, and no effort is taken to
manage complexity throughout the interface. Combined with an extremely low
amount of language visual richness, due to \ac{uml} standards, the downsides
of Visual Paradigm outweigh the good.

\paragraph{Visual Use Case} Unlike Visual Paradigm which supports the
creation of all \ac{uml} diagrams, Visual Use Case gives an in-depth
approach to the requirements workflow with a focus on use case creation.
Many different perspectives are provided as modularization tools to allow
the micromanagement of the different use cases. Some perspectives utilize
pieces of information from other perspectives for auto-completion, awarding
Visual Use Case with a very high interface efficiency as well as the
highest essential efficiency value. The extent at which the \ac{ide} goes
into detail does not make it ideal for beginners, however, and very few
popular \ac{ide} features are integrated in the design. Its high degree of
visual richness stacks up rather poorly as well, and the syntax within the
various perspectives is not enforced much. Overall, if Visual Use Case's
users can get it past its lack of accessibility, it turns out to be a very
powerful and in-depth tool for managing use cases.

\paragraph{WebRatio} Similar to Tersus, WebRatio provides a web application
creation interface with many popular \ac{ide} features. The overall visual
richness of WebRatio is improved over that of Tersus, however the essential
efficiency is slightly worse. WebRatio possesses a much larger amount of
clutter as well, but it also incorporates a couple more perspectives for
model management. Additionally similar to Tersus is the use of a
hierarchical design to manage complexity within the system, although the
syntax of the modeling language is not enforced either. Though WebRatio and
Tersus tend to be similar in many ways, each have their own pros and cons
which leaves it up for the user to determine his own preference of the
two.

\paragraph{YAWL} YAWL is a business workflow system that supports a simple,
\acs{uml}-like modeling language. Despite its simplicity, YAWL is targeted
toward skilled members of its domain. YAWL is lacking in its support of
popular \ac{ide} features and any form of active syntax checking. Complex
diagrams are also difficult to manipulate given the lack of complexity
management paradigms.  YAWL also only supports a slightly lower than
average number of visual richness variables in both the chrome and
language. The language itself especially suffers from lack of visual
richness with highly visually similar, square elements. However, YAWL's
interface has a very low clutter value and high degrees of efficiency.
Ultimately, the negatives outweigh the positives and, though minimal and
efficient, YAWL is overall not visually appealing and poor in \ac{ide} features.

%\begin{landscape}
%\begin{table*}
	%\centering
		%{\scriptsize
		%\begin{tabular}{l|cc|cccccccc|ccccc|ccc|ccc}
		  %\hline\hline
		  %&&&&&&&&&&&&&&&&&&&&&		\\[-1.5ex]
      %\textbf{IDE}			& \textbf{\scshape Audience}
                        %&
                        %& \textbf{Chrome}
                        %&&&&&&&
                        %& \textbf{Human Interface}
                        %&&&&
                        %& \textbf{Integration}
                        %&&
                        %& \textbf{Language Syntax}
                        %&&  \\[1ex]
                        %& \textbf{\scshape Domain}
                        %& \textbf{\scshape Skill}
                        %& \textbf{\scshape Features}
                        %& \textbf{\scshape Context}
                        %& \textbf{\scshape ToolRichness}
                        %& \textbf{\scshape Perspectives}
                        %& \textbf{\scshape Properties}
                        %& \textbf{\scshape Searchable}
                        %& \textbf{\scshape ToolStyle}
                        %& \textbf{\scshape Clutter}
                        %& \textbf{\scshape EEfficiency}
                        %& \textbf{\scshape IEfficiency}
                        %& \textbf{\scshape Keyboard}
                        %& \textbf{\scshape Mode}
                        %& \textbf{\scshape Devices}
                        %& \textbf{\scshape Relations}
                        %& \textbf{\scshape Output}
                        %& \textbf{\scshape Syntax}
                        %& \textbf{\scshape Complexity}
                        %& \textbf{\scshape Connection}
                        %& \textbf{\scshape LanguageRichness}			\\[1ex]
		  %\hline
		  %&&&&&&&&&&&&&&&&&&&&&		\\[-1.5ex]
      %\textbf{Alice3}   & Animation
                        %& Novice
                        %& 8
                        %& Yes
                        %& 6
                        %& 2
                        %& Omnipresent
                        %& No
                        %& Tabs
                        %& 3.40
                        %& 0.46
                        %& 0.08
                        %& Optional
                        %& Drag n Drop (1:1)
                        %& None
                        %& No
                        %& Direct Triggered
                        %& Implicit
                        %& Modularization
                        %& Overlapping Regions
                        %& 8			\\[1ex]
      %\textbf{AToMPM}	& Modeling
	%& Intermediate
	%& 4
	%& No
	%& 5
	%& 1
	%& Manual
	%& No
	%& Modular
	%& 2.60
	%& 0.23
	%& 0.15
	%& Optional
	%& Point n Click (1:n)
	%& None
	%& No
	%& Indirect Triggered
	%& Explicit
	%& None
	%& Linked Regions
	%& 6			\\[1ex]
      %\textbf{AudioMulch}	& Music
	%& Expert
	%& 7
	%& Yes
	%& 6
	%& 2
	%& Manual
	%& No
	%& Tree, Windows
	%& 3.13
	%& 0.31
	%& 0.31
	%& Simple
	%& Drag n Drop (1:1)
	%& Keyboards
	%& Yes
	%& Indirect Live
	%& Implicit
	%& None
	%& Linked Points
	%& 3			\\[1ex]
      %\textbf{Blender}	& 3D Modeling
	%& Expert
	%& 7
	%& Yes
	%& 7
	%& Multiple
	%& Omnipresent
	%& No
	%& Multiple
	%& 4.00
	%& 0
	%& -0.2
	%& Required
	%& Point n Click (1:1)
	%& None
	%& No
	%& Direct Live
	%& Implicit
	%& Modularization
	%& Geometric
	%& 9			\\[1ex]
      %\textbf{Cameleon}	& Prototyping
	%& Intermediate
	%& 3
	%& Yes
	%& 5
	%& 1
	%& Manual
	%& Yes
	%& Tree
	%& 2.07
	%& 0.23
	%& 0.31
	%& Required
	%& Drag n Drop (1:1)
	%& None
	%& Yes
	%& Direct Live
	%& Explicit
	%& Hierarchy
	%& Linked Points
	%& 5			\\[1ex]
      %\textbf{EMF}	& Software
	%& Expert
	%& 9
	%& Yes
	%& 5
	%& Multiple
	%& Manual
	%& Yes
	%& Tree, Icons
	%& 4.07
	%& 0.23
	%& 0.62
	%& Optional
	%& Point n Click (1:1)
	%& None
	%& No
	%& Indirect Triggered
	%& Explicit
	%& Modularization
	%& Linked Points, Linked Regions
	%& 5			\\[1ex]
      %\textbf{GNU Radio Companion}	& Software
	%& Intermediate
	%& 6
	%& No
	%& 4
	%& 1
	%& Manual
	%& Yes
	%& Icons
	%& 2.80
	%& 0.38
	%& 0.62
	%& Optional
	%& Drag n Drop (1:1)
	%& None
	%& No
	%& Indirect Triggered
	%& Explicit
	%& None
	%& Linked Points
	%& 2			\\[1ex]
      %\textbf{Grasshopper 3D}	& 3D Modeling
	%& Intermediate
	%& 7
	%& Yes
	%& 4
	%& 1
	%& None
	%& Yes
	%& Ribbons
	%& 2.80
	%& 0.31
	%& 0.31
	%& Optional
	%& Point n Click (1:1)
	%& None
	%& No
	%& Indirect Live
	%& Explicit
	%& None
	%& Linked Points
	%& 6			\\[1ex]
      %\textbf{Max}	& Music
	%& Expert
	%& 7
	%& No
	%& 4
	%& 2
	%& Manual
	%& Yes
	%& Tree, Windows
	%& 2.73
	%& 0.15
	%& 0.15
	%& Simple
	%& Drag n Drop (1:1)
	%& Microphones
	%& No
	%& Indirect Live
	%& None
	%& Modularization
	%& Linked Points
	%& 5			\\[1ex]
      %\textbf{MetaEdit+}	& Modeling
	%& Intermediate
	%& 7
	%& Yes
	%& 4
	%& 3
	%& Manual
	%& No
	%& Icons, Windows
	%& 1.93
	%& 0.31
	%& -0.08
	%& Simple
	%& Point n Click (1:1)
	%& None
	%& No
	%& Direct Live
	%& Implicit
	%& Modularization
	%& Linked Regions
	%& 5			\\[1ex]
      %\textbf{MIT AppInventor2}	& Software
	%& Novice
	%& 3
	%& Yes
	%& 4
	%& 2
	%& Omnipresent
	%& No
	%& Drawers
	%& 3.27
	%& 0.46
	%& 0.54
	%& Simple
	%& Drag n Drop (1:1)
	%& Mobile
	%& No
	%& Direct Triggered
	%& Implicit
	%& Modularization
	%& Overlapping Regions
	%& 9			\\[1ex]
      %\textbf{MST}	& Simulation
	%& Intermediate
	%& 6
	%& No
	%& 3
	%& 1
	%& Manual
	%& No
	%& Icons
	%& 2.40
	%& 0.38
	%& 0.38
	%& Simple
	%& Point n Click (1:1)
	%& None
	%& No
	%& Direct Live
	%& None
	%& None
	%& Overlapping Points
	%& 4			\\[1ex]
      %\textbf{Piet Creator}	& Software
	%& General
	%& 4
	%& No
	%& 7
	%& 1
	%& None
	%& No
	%& Icons
	%& 2.13
	%& 0.17
	%& -0.33
	%& None
	%& Point n Click (1:n)
	%& None
	%& No
	%& Indirect Triggered
	%& None
	%& None
	%& Geometric
	%& 4			\\[1ex]
      %\textbf{Scratch}	& Software
	%& Novice
	%& 3
	%& No
	%& 6
	%& 2
	%& None
	%& No
	%& Drawers
	%& 3.53
	%& 0.46
	%& 0.54
	%& Simple
	%& Drag n Drop (1:1)
	%& None
	%& No
	%& Direct Triggered
	%& Implicit
	%& Modularization
	%& Overlapping Regions
	%& 10			\\[1ex]
      %\textbf{SimuLink}	& Simulation
	%& Expert
	%& 7
	%& No
	%& 3
	%& 1
	%& Manual
	%& Yes
	%& Icons, Tree
	%& 3.80
	%& 0.31
	%& 0.15
	%& SImple
	%& Drag n Drop (1:1)
	%& None
	%& No
	%& Indirect Live
	%& None
	%& None
	%& Linked Regions
	%& 4			\\[1ex]
      %\textbf{Stencyl}	& Software
	%& Novice
	%& 6
	%& Yes
	%& 8
	%& Multiple
	%& Manual
	%& Yes
	%& Tabs, Icons
	%& 2.73
	%& 0.46
	%& 0.54
	%& Simple
	%& Drag n Drop (1:1)
	%& Mobile
	%& No
	%& Direct Triggered
	%& Implicit
	%& Modularization
	%& Overlapping Regions
	%& 10			\\[1ex]
      %\textbf{Tersus}	& Software
	%& Intermediate
	%& 8
	%& No
	%& 3
	%& 1
	%& Omnipresent
	%& No
	%& Drawers
	%& 3.20
	%& 0.31
	%& 0.15
	%& Simple
	%& Point n Click (1:1)
	%& None
	%& No
	%& Direct Triggered
	%& None
	%& Hierarchy
	%& Linked Points, Linked Regions, Geometric
	%& 3			\\[1ex]
      %\textbf{TouchDevelop}	& Software
	%& Novice
	%& 6
	%& Yes
	%& 6
	%& 1
	%& None
	%& No
	%& Icons
	%& 3.60
	%& 0.44
	%& 0.33
	%& None
	%& Point n Click (1:1)
	%& Mobile
	%& No
	%& Indirect Triggered
	%& Explicit
	%& Modularization
	%& Geometric
	%& 4			\\[1ex]
      %\textbf{UMLet}	& Modeling
	%& General
	%& 6
	%& No
	%& 3
	%& 1
	%& Omnipresent
	%& No
	%& Palette
	%& 3.00
	%& 0.31
	%& -0.15
	%& Simple
	%& Drag n Drop (1:1)
	%& None
	%& No
	%& Direct Live
	%& None
	%& None
	%& Overlapping Points
	%& 4			\\[1ex]
      %\textbf{VioletUML}	& Modeling
	%& Intermediate
	%& 6
	%& No
	%& 4
	%& 1
	%& Manual
	%& No
	%& Drawers
	%& 2.47
	%& 0.15
	%& 0.15
	%& Simple
	%& Point n Click (1:n)
	%& None
	%& No
	%& Direct Live
	%& Explicit
	%& None
	%& Linked Regions
	%& 4			\\[1ex]
      %\textbf{VisSim}	& Simulation
	%& Expert
	%& 8
	%& No
	%& 4
	%& 1
	%& Manual
	%& Yes
	%& Tree, Menu
	%& 3.33
	%& 0.31
	%& 0.31
	%& Simple
	%& Drag n Drop (1:1)
	%& None
	%& No
	%& Direct Live
	%& None
	%& Hierarchy
	%& Linked Points
	%& 4			\\[1ex]
      %\textbf{Visual Paradigm}	& Modeling
	%& Intermediate
	%& 7
	%& Yes
	%& 5
	%& 1
	%& Omnipresent
	%& No
	%& Icons, Menu
	%& 3.67
	%& 0.23
	%& 0.23
	%& Simple
	%& Drag n Drop (1:1)
	%& None
	%& No
	%& Direct Live
	%& Implicit
	%& None
	%& Linked Regions
	%& 2			\\[1ex]
      %\textbf{Visual Use Case}	& Modeling
	%& Intermediate
	%& 3
	%& No
	%& 4
	%& 7
	%& Manual
	%& No
	%& Icons, Tree
	%& 2.67
	%& 0.54
	%& 0.38
	%& Simple
	%& Point n Click (1:1)
	%& None
	%& No
	%& Direct Live
	%& None
	%& Modularization
	%& Linked Regions
	%& 3			\\[1ex]
      %\textbf{WebRatio}	& Software
	%& Intermediate
	%& 8
	%& No
	%& 4
	%& 3
	%& Omnipresent
	%& No
	%& Icons, Menu
	%& 3.93
	%& 0.23
	%& 0.23
	%& Simple
	%& Point n Click (1:1)
	%& None
	%& No
	%& Indirect Triggered
	%& None
	%& Hierarchy
	%& Linked Regions, Geometric
	%& 4			\\[1ex]
      %\textbf{YAWL}	& Workflow
	%& Expert
	%& 5
	%& No
	%& 5
	%& 2
	%& Omnipresent
	%& No
	%& Icons
	%& 2.27
	%& 0.31
	%& 0.31
	%& Simple
	%& Point n Click (1:n)
	%& None
	%& No
	%& Direct Triggered
	%& None
	%& None
	%& Linked Regions
	%& 4			\\[1ex]
			%\hline
		%\end{tabular}}
	%\caption{Feature matrix for all \acp{ide}.}
	%\label{tab:tools}
%\end{table*}
%\end{landscape}
